---
title: "FGEM_Simulations"
author: "Nicholas Knoblauch"
date: "October 12, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Simulating data

The data that give rise to the Bayes factors for this simulation come from the following scenario.  

Let $G$ be the number of genes, and $N$ be the number of samples,let $X_{i,j}$ denote a bernoulli random variable corresponding to some event (maybe a LoF de novo mutation) in gene $i$ in sample $j$, and let $X_i$ be a binomially distributed variable corresponding to the total number of events observed in gene $i$ (i.e $X_i \sim \text{Bernoulli}(N,p_i)$). If $Z_i=1$, Then $p_i$ is distributed according to a Beta distribution with paramters $\alpha=\beta=0.5$.  If $Z_i=0$, $p_i$ is 0.5.  

## One causal (continuous) annotation

For the simplest scenario, there will be one annotation will be continuous and uniform on $-1,1$, and one constant "intercept" annotation. The values of $\beta_0$, and $\beta_1$ are chosen so that the prior mean is $\sim 0.05$, corresponding to about 1000 genes being causal. I simulated the data and estimated $\beta_0$, and $\beta_1$ 100 times, and plotted the histogram of $\beta_1$

```{r one_causal,echo=FALSE}
library(dplyr)
library(ggplot2)
results_df <- simfunc(ngenes=20000,
                      samplesize=1000,
                      iterations=100,
                      betavec=c(-3,1.5))
ggplot(results_df)+geom_histogram(aes(x=Beta))+geom_vline(xintercept=1.5,col="purple")+ggtitle("Distribution of Beta\n One continuous feature")

```


## Two continuous features, joint model

Same setup, but now there are two features

```{r}
library(foreach)
annofunc <- list("Intercept"=rbinom,
                 "a"=runif)
annofeat <- list("Intercept"=list(size=1,prob=1),
                 "a"=list(min=-1,max=1))
ngenes <- 20000
A <- gen_anno(annofunc,annofeat,20000)
betavec <- c(-3,10)
tprior <- gen_p(betavec,x = A)
betap <- c(0.1,0.1)

comp_res <- foreach(i=1:100,.combine = "bind_rows") %do% {
  gprior=mean(tprior)
  sample_data <- sim_dat(tprior,ngenes,samplesize=100,betap) %>% mutate(tmu=(gprior*XH1)/((gprior*XH1)+(1-gprior)*XH0),
                                                                        bmu=(gprior*BF)/((gprior*BF)+(1-gprior)))
  
  tfbeta <- coefficients(glm(tmu~A+0,data=sample_data,family=quasibinomial(link="logit")))
  bfbeta <- coefficients(glm(bmu~A+0,data=sample_data,family=quasibinomial(link="logit")))
  tretdf <- true_est(fBeta = tfbeta,A = A,XH1 = sample_data$XH1,XH0 = sample_data$XH0)
  bretdf <- FGEM(fBeta = tfbeta,feat_mat = A ,BF = sample_data$BF)
}

```



```{r}
ggplot(comp_res)+geom_histogram(aes(x=Intercept_b))+geom_vline(aes(xintercept = Intercept))
```





```{r}
ggplot(comp_res)+geom_histogram(aes(x=Beta_b))+geom_vline(aes(xintercept = Beta))
```

```{r}
ggplot(comp_res)+geom_histogram(aes(x=Beta_t))+geom_vline(aes(xintercept = Beta))
```





## One causal, one non-causal feature, joint model

```{r}


feat_names <- c("Intercept","Causal","Noncausal")
annofunc <- list(rbinom,
                 runif,
                 runif)
names(annofunc) <- feat_names
annofeat <- list(list(size=1,prob=1),
                 list(min=-1,max=1),
                 list(min=-1,max=1))
names(annofeat) <- feat_names
betavec <- c(-3,1.5,0)
betadf <- data_frame(feature=feat_names[-1],trueBeta=betavec[-1])
results_df <- simfunc(betavec=betavec,annofunc=annofunc,annofeat=annofeat)

results_df <- inner_join(betadf,results_df)
ggplot(results_df)+geom_histogram(aes(x=Beta))+geom_vline(aes(xintercept=trueBeta),col="purple")+facet_wrap(~feature)+ggtitle("Distribution of Beta\n One causal and one non-causal feature")



```



