---
title: "FGEM"
author: "Nicholas Knoblauch"
date: "October 7, 2015"
output: html_document
---

####Preprocessing
```{r}
#library for transforming data
library(reshape2)
#libraries for pulling GO terms
library(topGO)
library(org.Hs.eg.db)
#library for creating sparse matrices
library(Matrix)

#File containing Bayes Factors for all I genes
BayesFactorFile <- "~/Dropbox/BayesianDA/FGEM/Data/TADA_ASC_SSC_results_Dec23.csv"
#Destination file for annotation matrix A
AnnotationMatrixOutfile <- "~/Dropbox/BayesianDA/FGEM/DataGOmat.RDS"
BF.df <- read.table(BayesFactorFile,header=T,sep=",")

#Create a new dataframe with just the Bayes Factor and the Bayes Factors
allgenes <- data.frame(Gene=BF.df$Gene,B=BF.df$BF,stringsAsFactors=F)
rownames(allgenes) <- allgenes$Gene

#Create an index for each gene, this will correspond to the row entry of the annotation matrix
allgenes$Gi <- as.integer(factor(allgenes$Gene))
#Pull all Molecular Function  and Biological Process GO terms for genes we have BF for and melt them into data frames
GO.MF.df <- melt(annFUN.org("MF",feasibleGenes=BF.df$Gene,mapping="org.Hs.eg.db",ID="symbol"))
colnames(GO.MF.df) <- c("Gene","GOT")
GO.BP.df <- melt(annFUN.org("BP",feasibleGenes=BF.df$Gene,mapping="org.Hs.eg.db",ID="symbol"))
colnames(GO.BP.df) <- c("Gene","GOT")

#Concatenate The BP and MF dataframes, pull Gene indices from before and generate annotation indices
GO.df <- rbind(GO.MF.df,GO.BP.df)
GO.df$Gi <- allgenes[GO.df$Gene,"Gi"]
GO.df$GOi <- as.integer(factor(GO.df$GOT))

#Generate sparse matrix of annotations.  Specify row, column and value (1 in this case) for each non-zero entry
GOmat <- sparseMatrix(i=GO.df$Gi,j = GO.df$GOi,x = rep(1,nrow(GO.df)),
                       dimnames=list(as.character(allgenes$Gene),
                                     as.character(levels(factor(GO.df$GOT)))))
#About half (6368/13324) of GO terms have fewer than 3 genes with that annotation, remove these terms.  
GOmat <- GOmat[,colSums(GOmat)>=3]

#We now order the remaining terms by the significance of their association with the Bayes Factor
#(This takes a while, so it's commented out)
# B <- allgenes[rownames(GOmat),"B"]
# Betap<- numeric(ncol(GOmat))
# for(i in 1:ncol(GOmat)){
#     if(i%%100==0){
#         print(i)
#     }
#     Betap[i] <- summary(lm(B~GOmat[,i]))[["coefficients"]][2,4]
# }
# GOmat <- GOmat[,order(Betap)]
# saveRDS(GOmat,file=AnnotationMatrixOutfile)
```


####EM


```{r}
#Usually read from the command line to get which iteration to use
#args <- commandArgs(trailingOnly=T)
GOmatfile <- "~/Dropbox/BayesianDA/FGEM/Data/GOmat.RDS"
BayesFactorFile <- "~/Dropbox/BayesianDA/FGEM/Data/TADA_ASC_SSC_results_Dec23.csv"
Startterm <- 1
Numterms <- 10
EMiter <- 50
Outdir <- "~/Dropbox/BayesianDA/"


GOmat <- readRDS(GOmatfile)
BF.df <- read.table(BayesFactorFile,header=T,sep=",")

#To start the EM algorithm, we need a starting guess for Beta. To do this we will use 
#logistic regression on thresholded Bayes Factors (Bayes Factor>3) 
BF.df$sig <- ifelse(BF.df$BF>5,1,0)
Z <- BF.df$sig




allgenes <- data.frame(Gene=BF.df$Gene,Z,B=BF.df$BF,stringsAsFactors=F)
rownames(allgenes) <- allgenes$Gene

#Pull the relevant rows from the dataframe

B <- allgenes[rownames(GOmat),"B"]

####Starting EM


#If the number of terms to be tested is greater than the number of remaining terms
# we change the number of terms to be tested
if(Startterm+Numterms>ncol(GOmat)){
    Numterms <- ncol(GOmat)-Startterm
}

#Subset the GO matrix to only the terms we're going to test
GOmat <- GOmat[,Startterm:(Startterm+Numterms)]


#Computing the null model
likfun <- function(x,B){
  sum(log((1/(1+exp(-x)))*B+(1-(1/(1+exp(-x))))))
}
mo <- optimize(likfun,interval = c(-10,10),B=B,maximum=T)


#Function that performs EM
FGEM <- function(x,B,Z=NULL,iters=NULL,tol=NULL,keep=F,NullLogLikelihood=NULL){
  #x is a vector of length I with relevant annotations
  #B is a vector of length I with Bayes Factors
  #Z is a vector of length I with the initial guess for membership (1's or 0's)
  ## Z can be NULL, in which case Z will be randomly sampled from a binomial with probabilities according to the Bayes Factor.
  
  #iters is the number of EM iterations.  If not specified, EM will be run untill convergence specified by tol.
  
  if(is.null(NullLogLikelihood)){
    likfun <- function(x,B){
      sum(log((1/(1+exp(-x)))*B+(1-(1/(1+exp(-x))))))
    }
    NullLogLikelihood  <- optimize(likfun,interval = c(-10,10),B=B,maximum=T)$objective
  }
  
  if(is.null(tol)&is.null(iters)){
    stop("Must specify tol or iters")
  }
  if(ifelse(is.null(Z),
            length(x)!=length(B),
            (length(x)!=length(B))|(length(x)!=length(Z))|(length(B)!=length(Z)))){
    stop("Length of x and B (and Z, if specified) must be equal")
  }
  
  if(is.null(Z)){
    pB <- ecdf(B)(B)
    sigcount <- sum(pB)/sum(B>10)
    Z <- rbinom(length(B),size = 1,prob = pB/sigcount)
  }
  if(!is.null(iters)&is.null(tol)){
    tol <- 0
  }
  if(!is.null(tol)&is.null(iters)){
    iters <- 1000
  }
  if(keep){
    matdim <- iters
    Beta <- matrix(0,2,matdim)
    Beta[,1] <- coefficients(glm(Z~x,family=binomial))
    pvec <- 1/(1+exp(-(Beta[1,1]+Beta[2,1]*x)))
    LogLikelihood <- numeric(matdim)
    LogLikelihood[1] <- sum(log(pvec*B+(1-pvec)))
    i <- 2
    while ((i <= iters)&(i<1000)){
      print(i)
     #This is the mixture proportion
      uvec  <- (pvec*B)/((pvec*B)+(1-pvec))
      #This is where we optimize the Q function (in this case, using logistic regression)
      Beta[,i] <- coefficients(glm(uvec~x,family=quasibinomial(link="logit")))
      pvec  <- 1/(1+exp(-(Beta[1,i]+Beta[2,i]*x)))
      #If the difference between the current log likelihood and the previous log likelihood is below the tolerance, we quit
      rettol <- sum(log(pvec*B+(1-pvec)))-LogLikelihood[i-1]
      if(rettol<tol){
        LogLikelihood[i] <- sum(log(pvec*B+(1-pvec)))
        break
      }
      LogLikelihood[i] <- sum(log(pvec*B+(1-pvec)))
      #If we're about to stop and we still aren't better than the null model, we keep going (up to 1k iterations)
      if(i==(iters-1)){
        if(LogLikelihood[i]<NullLogLikelihood){
          oBeta <- Beta
          oLogLikelihood <- LogLikelihood
          Beta <- matrix(0,2,iters+10)
          Beta[,1:iters] <- oBeta
          LogLikelihood <- numeric(iters+50)
          LogLikelihood[1:iters] <- oLogLikelihood
          iters <- iters+10
        }
      }
      i <- i+1
    }
    
    Chisq <- -2*(NullLogLikelihood-LogLikelihood[matdim])
    retlist <- list(Intercept=Beta[1,matdim],Beta=Beta[2,matdim],
                    LogLikelihood=LogLikelihood[matdim],
                    NullLogLikelihood=NullLogLikelihood,
                    Chisq=Chisq,tol=rettol)
  }
  else{
    Beta <- coefficients(glm(Z~x,family=binomial))
    pvec <- 1/(1+exp(-(Beta[1]+Beta[2]*x)))
    LogLikelihood <- sum(log(pvec*B+(1-pvec)))
    i <- 1
    while ((i < iters)&(i<1000)){
      #This is the mixture proportion
      uvec  <- (pvec*B)/((pvec*B)+(1-pvec))
      #This is where we optimize the Q function (in this case, using logistic regression)
      Beta <- coefficients(glm(uvec~x,family=quasibinomial(link="logit")))
      pvec  <- 1/(1+exp(-(Beta[1]+Beta[2]*x)))
      #If the difference between the current log likelihood and the previous log likelihood is below the tolerance, we quit
      rettol <- sum(log(pvec*B+(1-pvec)))-LogLikelihood
      if(rettol<tol){
        LogLikelihood <- sum(log(pvec*B+(1-pvec)))
        break
      }
      LogLikelihood <- sum(log(pvec*B+(1-pvec)))
      #If we're about to stop and we still aren't better than the null model, we keep going (up to 1k iterations)
      if(i==(iters-1)&LogLikelihood<NullLogLikelihood){
        iters <- iters+10
      }
      i <- i+1
    }
    Chisq <- -2*(NullLogLikelihood-LogLikelihood)
    retlist <- list(Intercept=Beta[1],Beta=Beta[2],
                    LogLikelihood=LogLikelihood,
                    NullLogLikelihood=NullLogLikelihood,
                    Chisq=Chisq,tol=rettol)
  } 
  if(keep){
    mats <- list(Betas=Beta,LogLikelihood=LogLikelihood)
    return(list(retlist,mats))
  }else{
    return(retlist)
  }
}

```

Let's run it on five different features. By using the `keep=T` argument, we can show the convergence of Beta and the likelihood function.

```{r}

NullLogLikelihood <- mo$objective
Niter <- FGEM(x = GOmat[,10],B = B,iters = 50)
FiveIters <- apply(GOmat[,1:5],2,FGEM,B=B,iters=50,keep=F,NullLogLikelihood=NullLogLikelihood)
fi <- ldply(FiveIters,data.frame)

```


###Top Results

After running in parallel across 100 CPUs on pps, we can summarize the results

```{r}

GOmat <- readRDS("~/Dropbox/BayesianDA/GOmat.RDS")
CompletedAnalyses <- "~/Dropbox/BayesianDA/RDS_BDF/"
allBDFfiles <- dir(CompletedAnalyses,full.names = T)
BDFl <- lapply(allBDFfiles,readRDS)
allBDF <- ldply(allBDFfiles,readRDS)
allBDF <- allBDF[!is.na(allBDF$GO),]
allBDF$chisq <- -2*(NullLogLikelihood-log(allBDF$likalt))
badBDF <- allBDF[allBDF$chisq<0,]



allBDF <- allBDF[ allBDF$chisq>0,]
allBDF <- allBDF[!duplicated(allBDF$GO),]
rownames(allBDF) <- allBDF$GO
allBDF <- allBDF[order(allBDF$p),]
allBDF$padj <- p.adjust(allBDF$p,method = "fdr")

sigGO <- allBDF$GO[ allBDF$p<0.05]

sigmat <- GOmat[,which(colnames(GOmat) %in% sigGO)]

dsm <- data.frame(which(sigmat!=0,arr.ind = T),stringsAsFactors = F)
dsm$gene <- rownames(sigmat)[dsm[,1]]
dsm$GO <- colnames(sigmat)[dsm[,2]]
dsm$pval <- allBDF[dsm$GO,"p"]
GO.annotation <- select(GO.db,dsm$GO,c("TERM","ONTOLOGY"))
uGO <- GO.annotation[!duplicated(GO.annotation$GOID),]
rownames(uGO) <- uGO$GOID
allBDF$TERM <- uGO[ allBDF$GO,]
dsm <- cbind(dsm,GO.annotation)
dsm <- dsm[,-c(1,2)]


```

How do these p values compare to the fisher's exact test(with a cutuff of 5)
```{r}

FGEM.df <- readRDS("~/Dropbox/BayesianDA/FGEM.RDS")
FGEM.df <-rename(FGEM.df,replace = c(".id"="GOid"))
FGEM.df <- FGEM.df[!duplicated(FGEM.df$GOid),]

FGEM.df <- FGEM.df[order(FGEM.df$p),]
#Use BF 50
 Z <- ifelse(BF.df$BF>10,1,0)
 
afish <- apply(GOmat,2,function(x){
  fisher.test(table(x,Z),alternative = "greater")[["p.value"]]
})

fish.df <- data.frame(GOid=names(afish),fishp=afish,stringsAsFactors=F)
#saveRDS(fish.df,"~/Dropbox/BayesianDA/fish.df.RDS")
fish.df <-readRDS("~/Dropbox/BayesianDA/fish.df.RDS")
FGEM.df <- FGEM.df[FGEM.df$GOid %in% rownames(fish.df),]
fish.df <- fish.df[FGEM.df$GOid,]

plot(FGEM.df$p,fish.df$fishp,xlab="FGEM p",ylab="fisher exact p")

bmethod <- data.frame(GOid =rownames(fish.df),FGEMp=FGEM.df$p,fishp=fish.df$fishp,stringsAsFactors=F)

bmethod <- bmethod[order(bmethod$fishp),]
plot(bmethod$FGEMp[bmethod$fishp<0.05],bmethod$fishp[bmethod$fishp<0.05])
t.test(B~GOmat[,"GO:0032206"])
fisher.test(table(GOmat[,"GO:0032206"],Z))


Geneiter <- FGEM(x = avg.exp[,2],B = B,iters = 50,keep=F)

FGEM.sig.GO <- rownames(bmethod[bmethod$FGEMp<0.05,]),alternative = "greater")
fisher.sig.GO <- rownames(bmethod[bmethod$fishp<0.05,])

fisher.test(table(bmethod$FGEMp<0.05,bmethod$fishp<0.05),alternative="greater")
table(FGEM.sig.GO,fisher.sig.GO)
