---
title: "FGEM_Simulation"
author: "Nicholas Knoblauch"
date: "May 2, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Simulation

I will use the full suite of generated annotations, I will simulate effect sizes, then simulate data, and try to recapture the results

## Annotations

+ Gene Ontology
+ Exac 

### Gene Ontology

We'll start by reading in the data, the next step 

```{r libs,echo=FALSE}
library(rhdf5)
library(Matrix)
library(ggplot2)
library(SQUAREM)
source("~/Dropbox/BayesianDA/FGEM/FGEM_Func.R")
annof <- "~/Dropbox/BayesianDA/FGEM/Data/Annomat.h5"

datmat <- sparseAnnoread(annof,"Annotations/GO/GOmat",
                       rowpath="GeneList/GL",
                       colpath="Annotations/GO/GOterms")


ggplot()+geom_histogram(aes(x=log(colSums(datmat))))


```
Master simulation parameters


```{r}
Ncase <- 500
Nctrl <- 500

nfeat <- 100
gambar <- 5
sigma <- .1

testmat <- datmat[,sample(ncol(datmat),100,replace=F)]
testmat <- testmat[rowSums(testmat)>0,]
ngenes <- nrow(testmat)

```

### Simulation scenarios
#### 1) Dense effect size

For scenario 1, effect size will be modeled as non-sparse (dense).

Effect size will be additive, centered at 0, normally distributed, with variance 1.

```{r dense}
Betas <- rnorm(nfeat)

pr <- GenPrior(c(0,Betas),testmat)[,1]
effc <- (pr*mean(1-pr))/((1-pr)*mean(pr))



fB <- squarem(par=effc/(effc+exp(-effc)),fixptfn=Logit_FGEM,
           objfn=Logit_log_lik,x=as.matrix(testmat),B=effc,control=list(trace=T))
bB  <- squarem(par=effc/(effc+exp(-effc)),fixptfn=Logit_FGEM,
           objfn=Logit_log_lik,x=as.matrix(testmat),B=effc,control=list(trace=T))

frBetas <- apply(as.matrix(testmat),2,function(x){
  fB <- squarem(par=effc/(effc+exp(-effc)),fixptfn=Logit_FGEM,
           objfn=Logit_log_lik,x=x,B=effc,control=list(trace=T))
  fp <- fB$par[,1]
  fBeta <- coefficients(glm(fp~as.matrix(x),family=quasibinomial(link="logit")))
  return(fBeta[-1])
})

bBetas <- apply(as.matrix(testmat),2,function(x){
  fB <- squarem(par=effc/(effc+exp(-effc)),fixptfn=Logit_FGEM,
           objfn=Logit_log_lik,x=x,B=effc,control=list(trace=T))
  fp <- fB$par[,1]
  fBeta <- coefficients(glm(fp~as.matrix(x),family=quasibinomial(link="logit")))
  return(fBeta[-1])
})

hist(rbeta(1000,shape1=16,shape2 = 24))

head(percent_rank(effc))
npar <- qbeta(percent_rank(effc),shape1 = 16,shape2 = 24)

bB  <- squarem(par=npar,fixptfn=Logit_FGEM,
           objfn=Logit_log_lik,x=as.matrix(testmat),B=effc,control=list(trace=T))

plot(bp,pr)


bp <- bB$par[,1]

bBeta <- coefficients(bayesglm(bp~as.matrix(testmat),family=quasibinomial(link="logit")))
plot(bBeta[-1],Betas)
cor(bBeta[-c(1,which.min(bBeta))],Betas[-which.min(bBeta)])
boxplot(fBeta[-1]~caus)
plot(fBeta[-1],Betas)
table(round(fp,0),caus)
effc/(effc+exp(-effc))
head(fB)

par <- effc/effc+exp(-effc)




